---
title: Plots in R
description: "The basics of plots in R"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(testwhat)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
tutorial_options(exercise.checker = testwhat_learnr)

df <- POL306::state_cig

```

## A Note on Data

Throughout this tutorial we will use a dataset that looks at healthcare spending, cigarette taxes, and a variety of political variables. The unit of analysis for it is the state-year, and the data comes from 2001 to 2014. The variables are:

- `cig_tax` is the tax (in dollars) on a packet of cigarettes. 
- `female_leg` is the percentage of women in the state legislature.
- `liberal` is the percentage of voters that identify as liberal.
- `dem_gov` is an indicator that is 1 if the governor is a Democrat and 0 if they are not.
- `healthspendpc` is the amount spent perperson on healthcare. 
- `state` is the name of the state
- `year` is the year. 


## Histogram

First we will briefly explore histograms in R. 

Histograms are done by using the function `hist()`. Like most of our functions it expects that the first thing it encounters will be the variable you want a histogram of. It also has a lot of other attributes:

- `xlab=` or `ylab=`: Will change the label used for the y or x axis. Be sure to put whatever you want it to be in quotes: `xlab="Unemployment Rate"`
- `main=`: Will change the main title in a similar way to `xlab` or `ylab`
- `break = ...:` Will set the way the data is binned. Can take a lot of options, most common though is the number of bins you want. To make your histogram have 5 boxes you set break=5

**Instructions**:

Make a histogram of `female_leg` and change the xaxis and the main title.



```{r histogram, exercise=TRUE}
names(df)
```

```{r histogram-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

names(df)

hist(df$female_leg, xlab="Proportion Women in Legislature", main="Histogram of Women in State-Leg") 
```

```{r histogram-check, eval=FALSE}
ex() %>% check_function("hist") %>% {
  check_arg(., "xlab")
  check_arg(., "main")
}
```

## Density Plots

Now we will explore density plots in R. 

Density plots require using two functions at once `density()` and `plot()`. You can nest them so that they happen at the same time like: `plot(density(x))` where x is what you want to plot the the density of. You can change attributes of the plot using the same attributes from the `hist()` function as long as they are outside the first set of parentheses: `plot(density(x), xlab="X Label")`. 

You can see how different weighting schemes impact the density plot by setting `adjust=` to something. The default would be: `plot(density(x, adjust=1))`.

**Instructions**:

Make a density plot of `female_leg` and change the `adjust` argument to something other than 1.



```{r density, exercise=TRUE}
names(df)
```

```{r density-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

names(df)

plot(density(df$female_leg, adjust=.5, na.rm=T))
```

```{r density-check, eval=FALSE}
ex() %>% check_function("density") %>% {
  check_arg(., "adjust")
 }
```


## Scatter Plots

Plots in R can be highly customized. Scatter plots all use the same function though: `plot()`. `plot()` expects one or two vectors that it will plot. If you want to ensure which one is on the y axis, and which is on the x axis you can explicitly identify them: `plot(y=, x=)`. Beyond this there are additional arguments you can call:

- `xlab=` or `ylab=`: Will change the label used for the y or x axis. Be sure to put whatever you want it to be in quotes: `xlab="Unemployment Rate"`
- `main=`: Will change the main title in a similar way to `xlab` or `ylab`
- `pch=`: Can be used to changes the symbol used for points. You can set it to any integer between 0 and 25. For example `pch=2` makes the symbols into open triangles. The default is `pch=1`. 
- `col=`: Will set the color used for the points. You can pick from a lot of basic colors like: red, green, blue, yellow, orange, etc. Be sure to put the color name in quotes. 
- `type=`: Will change the type of plot. The default `type="p"` will make a scatter plot, while `type="l"` will make a line plot. 

Putting this all together would look something like:
```
plot(x=df$var1, y=df$var2, pch=2, col="red", main="Var 2 vs Var 1", xlab="Var 1",ylab="Var 2") 
```

**Instructions**:

Make a plot between `liberal` and `cig_tax`. Change `pch`, `col`, and `main`.



```{r scatter-plots, exercise=TRUE}
names(df) 
```

```{r scatter-plots-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

names(df) 

plot(x=df$liberal, y=df$cig_tax, pch=2, col="red", main="Liberal vs Cigarette Tax")
```

```{r scatter-plots-check, eval=FALSE}
ex() %>% {
  check_function(., "plot" ) %>% {
    check_arg(., "pch")
    check_arg(., "col")
    check_arg(., "main")
  } 
}
```
