---
title: Bivariate Linear Regression
description: ""
output:
 learnr::tutorial:
   theme: sandstone
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(testwhat)
library(POL306)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
tutorial_options(exercise.checker = testwhat::testwhat_learnr)

df <- state_cig_df


```

## A Note on Data

Throughout this tutorial we will use a dataset that looks at healthcare spending, cigarette taxes, and a variety of political variables. The unit of analysis for it is the state-year, and the data comes from 2001 to 2014. The variables are:

- `cig_tax` is the tax (in dollars) on a packet of cigarettes. 
- `female_leg` is the percentage of women in the state legislature.
- `liberal` is the percentage of voters that identify as liberal.
- `dem_gov` is an indicator that is 1 if the governor is a Democrat and 0 if they are not.
- `healthspendpc` is the amount spent perperson on healthcare. 
- `state` is the name of the state
- `year` is the year. 


## Histogram

First we will briefly explore histograms in R. 

Histograms are done by using the function `hist()`. Like most of our functions it expects that the first thing it encounters will be the variable you want a histogram of. It also has a lot of other attributes:

- `xlab=` or `ylab=`: Will change the label used for the y or x axis. Be sure to put whatever you want it to be in quotes: `xlab="Unemployment Rate"`
- `main=`: Will change the main title in a similar way to `xlab` or `ylab`
- `break = ...:` Will set the way the data is binned. Can take a lot of options, most common though is the number of bins you want. To make your histogram have 5 boxes you set break=5

**Instructions**:

Make a histogram of `female_leg` and change the xaxis and the main title.



```{r histogram, exercise=TRUE}
names(df)
```

```{r histogram-solution, eval=FALSE}
df <- POL306::state_cig_df

names(df)

hist(df$female_leg, xlab="Proportion Women in Legislature", main="Histogram of Women in State-Leg") 
```

```{r histogram-check, eval=FALSE}
ex() %>% check_function("hist") %>% {
  check_arg(., "xlab")
  check_arg(., "main")
}
```

## Density Plots

Now we will explore density plots in R. 

Density plots require using two functions at once `density()` and `plot()`. You can nest them so that they happen at the same time like: `plot(density(x))` where x is what you want to plot the the density of. You can change attributes of the plot using the same attributes from the `hist()` function as long as they are outside the first set of parentheses: `plot(density(x), xlab="X Label")`. 

You can see how different weighting schemes impact the density plot by setting `adjust=` to something. The default would be: `plot(density(x, adjust=1))`.

**Instructions**:

Make a density plot of `female_leg` and change the `adjust` argument to something other than 1.



```{r density, exercise=TRUE}
names(df)
```

```{r density-solution, eval=FALSE}
df <- POL306::state_cig_df

names(df)

plot(density(df$female_leg, adjust=.5))
```

```{r density-check, eval=FALSE}
ex() %>% check_function("density") %>% {
  check_arg(., "adjust")
 }
```


## Scatter Plots

Plots in R can be highly customized. Scatter plots all use the same function though: `plot()`. `plot()` expects one or two vectors that it will plot. If you want to ensure which one is on the y axis, and which is on the x axis you can explicitly identify them: `plot(y=, x=)`. Beyond this there are additional arguments you can call:

- `xlab=` or `ylab=`: Will change the label used for the y or x axis. Be sure to put whatever you want it to be in quotes: `xlab="Unemployment Rate"`
- `main=`: Will change the main title in a similar way to `xlab` or `ylab`
- `pch=`: Can be used to changes the symbol used for points. You can set it to any integer between 0 and 25. For example `pch=2` makes the symbols into open triangles. The default is `pch=1`. 
- `col=`: Will set the color used for the points. You can pick from a lot of basic colors like: red, green, blue, yellow, orange, etc. Be sure to put the color name in quotes. 
- `type=`: Will change the type of plot. The default `type="p"` will make a scatter plot, while `type="l"` will make a line plot. 

Putting this all together would look something like:
```
plot(x=df$var1, y=df$var2, pch=2, col="red", main="Var 2 vs Var 1", xlab="Var 1",ylab="Var 2") 
```

**Instructions**:

Make a plot between `liberal` and `cig_tax`. Change `pch`, `col`, and `main`.



```{r scatter-plots, exercise=TRUE}
names(df) 
```

```{r scatter-plots-solution, eval=FALSE}
df <- POL306::state_cig_df

names(df) 

plot(x=df$liberal, y=df$cig_tax, pch=2, col="red", main="Liberal vs Cigarette Tax")
```

```{r scatter-plots-check, eval=FALSE}
ex() %>% {
  check_function(., "plot" ) %>% {
    check_arg(., "pch")
    check_arg(., "col")
    check_arg(., "main")
  } 
}
```


## Bivariate Linear Regression

The most important function you will use for estimating linear regression models is `lm()`. This will expect a formula and will output something called a model object that you will want to save so you can then manipulate it. 

For example, if you have a dataframe called `data.frame` and three variables named: `var1` and `var2`. If you want to estimate a linear regression with `var1` as the dependent variable and `var2` as the independent variable you would write:

```
mod <- lm(var1 ~ var2, data=data.frame)
```

This won't output anything but you can look at the results using `summary()` like this:

`summary(mod)`

**Instructions**:

We will start by just estimating a linear regression of the cigarette tax (the DV) as a function of the percent of liberals in a state. The variables are all in the `df` dataset and are: `cig_tax`, and `liberal`. 

You should save the linear model as `mod` (like in the example above) and then output the results using the `summary()` function.



```{r linear, exercise=TRUE}
names(df) 
```

```{r linear-solution, eval=FALSE}
df <- POL306::state_cig_df

names(df)

mod <- lm(cig_tax~liberal, data=df)
summary(mod)
```

```{r linear-check, eval=FALSE}
ex() %>% {
  check_function(., "lm" ) %>%
    check_arg("formula") %>% check_equal()
  check_function(., "summary")
}
```



## Linear Regression Extended

The reason that we save the `lm()` object is that there are a lot of things we can do with it after we are done. Here are a few things:

- `coef(mod)` will create a vector of just the coefficients, which can be useful if we want to do something to them. 
- `mod$residuals` is a vector of all the residuals for the regression. We could easily plot them with `plot(mod$residuals)`
- `nobs(mod)` calculates the number of observations in the final dataset.
- `mod$fitted.values` is a vector of the predictions from the regression.

**Instructions**:

Using the `mod` we made in the last assignment (I included the code here), do the following:

1. Get the number of the observations included in the model. 
2. Plot the residuals.


```{r linear-2, exercise=TRUE}
names(df)

mod <- lm(cig_tax~liberal, data=df)
summary(mod)
```

```{r linear-2-solution, eval=FALSE}
df <- POL306::state_cig_df

names(df)

mod <- lm(cig_tax~liberal, data=df)
summary(mod)

nobs(mod)
plot(mod$residuals)
```

```{r linear-2-check, eval=FALSE}
ex() %>% {
  check_function(., "plot")
  check_function(., "nobs" )
}

```


## Plotting a Regression Line

You can add the regression line to a scatter plot by using the `abline()` function after you make a scatter plot. You just have to put the model object (`mod`) into the `abline()` function. I've recreated the regression you looked at before with a scatter plot. Now you have to add a regression line. 


```{r linear-plot, exercise=TRUE}
names(df)

mod <- lm(cig_tax~liberal, data=df)
plot(y=df$cig_tax, x=df$liberal)
```

```{r linear-plot-solution, eval=FALSE}
df <- POL306::state_cig_df

names(df)

mod <- lm(cig_tax~liberal, data=df)
plot(y=df$cig_tax, x=df$liberal)
abline(mod)
```

```{r linear-plot-check, eval=FALSE}
ex() %>% 
  check_function("abline")

```


