---
title: Plots in R
description: "The basics of plots in R"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(testwhat)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
tutorial_options(exercise.checker = testwhat_learnr)

df <- POL306::state_cig

```

## A Note on Data

Throughout this tutorial we will use a dataset that looks at healthcare spending, cigarette taxes, and a variety of political variables. The unit of analysis for it is the state-year, and the data comes from 2001 to 2014. The variables are:

- `cig_tax` is the tax (in dollars) on a packet of cigarettes. 
- `female_leg` is the percentage of women in the state legislature.
- `liberal` is the percentage of voters that identify as liberal.
- `dem_gov` is an indicator that is 1 if the governor is a Democrat and 0 if they are not.
- `healthspendpc` is the amount spent perperson on healthcare. 
- `state` is the name of the state
- `year` is the year. 

## ggplot 

In this exercise we are going to use the package [ggplot2](https://ggplot2.tidyverse.org/) to make plots. This library is widly used to make plots and has a lot of customizability that makes it powerful. It also uses a specific way of creating plots that involves adding together functions.

Each ggplot2 call though starts with the function `ggplot()` and includes the data you are going to use in the `data=` attribute. 

**Instructions:**

Start now by running loading the ggplot2 library `library(ggplot2)` and running the initial function `ggplot(data=df)` (df is the data we are using here). 


```{r ggplot, exercise=TRUE}
```

```{r ggplot-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this
library(ggplot2)
ggplot(data=df)
```

```{r ggplot-check, eval=FALSE}
ex() %>% check_function("ggplot") %>% {
  check_arg(., "data")
}
```


## Histogram

You successfully made a grey box! Now lets start filling that grey box with a plot. You add `geom_*()` functions to build your plot, and in them you tell R what variables to use. You do that by mapping aesthetics to different variables in your data. 

For example, a histogram is `geom_histogram()` and if you want it to make a histogram of the cigarette tax on the x axis you'd write: `geom_histogram(mapping=aes(x=cig_tax))`. Note that we are assigning an aesthetics (`aes()`) to a `mapping=` for `geom_histogram()`. 

Putting this together with what we have before is: 

```
ggplot(data=df) + geom_histogram(mapping=aes(x=cig_tax))
```

**Instructions**:

Create a histogram of the percentage of female legislators in a state (`female_leg`)



```{r histogram, exercise=TRUE}
library(ggplot2)
```

```{r histogram-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

library(ggplot2)
ggplot(data=df) + geom_histogram(mapping=aes(x=female_leg))
 
```

```{r histogram-check, eval=FALSE}

ex() %>% {
  check_function(., "ggplot") %>% check_arg("data") %>% check_equal()
  check_function(., "aes") %>% {
    check_arg(., "x") %>% check_equal(eval = FALSE) 
  }
  check_function(., "geom_histogram")
}
```

## Density Plots

The powerful thing with `ggplot2` is that you can easily switch the way your data is displayed by changing the `geom_*()` function. Along with `geom_histogram()` there is an `geom_density()` that creates a density plot. 

**Instructions**:

Switch your histogram of percentage of female legislators into a density plot. 


```{r density, exercise=TRUE}
library(ggplot2)
```

```{r density-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

library(ggplot2)
ggplot(data=df) + geom_density(mapping=aes(x=female_leg))

```

```{r density-check, eval=FALSE}
ex() %>% {
  check_function(., "ggplot") %>% check_arg("data") %>% check_equal()
  check_function(., "aes") %>% {
    check_arg(., "x") %>% check_equal(eval = FALSE) 
  }
  check_function(., "geom_density")
}
```

## Customizing Labels and Titles

To create titles and labels for your plot you use the `labs()` function which takes arguments for the x axis (`x=`), the y axis (`y=`) and the main title (`title=`). For example to set the main title you would add on `labs(title="Histogram of Variable")`. Be sure to put the language you want to use in quotations, and separate each argument with a comma (if you use multiple). 

**Instructions**:

Provide an x and y label and a title for the histogram of female legislators. B

```{r label, exercise=TRUE}
library(ggplot2)
```

```{r label-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

library(ggplot2)
ggplot(data=df) + geom_histogram(mapping=aes(x=female_leg)) + 
  labs(y="Frequency", x="Percent Female", title="Histogram of Percent Female Legislators")

```

```{r label-check, eval=FALSE}
ex() %>% {
  check_function(., "ggplot") %>% check_arg("data") %>% check_equal()
  check_function(., "aes") %>% {
    check_arg(., "x") %>% check_equal(eval = FALSE) 
  }
  check_function(., "geom_histogram") 
  check_function(., "labs") %>%{
    check_arg(., "x")
    check_arg(., "y")
    check_arg(., "title")
  }
}
```

## Scatter Plots

Making a scatter plot in `ggplot2` is relatively easy as well, you just need to switch to `geom_point()` (because you are making points) and add a y variable along with your x variable (something like `aes(x=var1, y=var2))`. 

**Instructions**:

Make a scatter plot with `cig_tax` on the y axis and `female_leg` on the x. Be sure to set labels for the x and y axis along with a title for the plot. 


```{r scatter-plots, exercise=TRUE}
library(ggplot2)
```

```{r scatter-plots-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

ggplot(data=df) + geom_point(mapping=aes(x=female_leg, y=cig_tax)) + 
  labs(y="Cigarette Tax", x="Percent Female", 
       title="Cig Tax vs Female Legislature")

```

```{r scatter-plots-check, eval=FALSE}
ex() %>% {
  check_function(., "ggplot") %>% check_arg("data") %>% check_equal()
  check_function(., "geom_point")
  check_function(., "aes") %>% {
    check_arg(., "x") %>% check_equal(eval = FALSE) 
    check_arg(., "y") %>% check_equal(eval = FALSE) 
  }
  check_function(., "labs") %>%{
    check_arg(., "x")
    check_arg(., "y")
    check_arg(., "title")
  }
}
```




## Other Aesthetics 

ggplot has a wide variety of ways to customize a plot, and these can be used to either show more information about your observations _or_ just to make things look pretty.

For example you can change the colors of the points using the `color=` argument. If you set this in the `aes()` function and assign it to a variable then it will color the points based on that variable. You could also just set it outside the `aes()` function but inside the `geom_point()` function to any color like `color="green"` to set all the points to green. Concretely the difference looks like: 

```
### Color the dots based on how liberal a state is
ggplot(data=df) + 
  geom_point(mapping=aes(x=female_leg, y=cig_tax, 
                          color=liberal)) 
```

```
### Color all the dots red
ggplot(data=df) + 
  geom_point(mapping=aes(x=female_leg, y=cig_tax), 
                          color="red") 
```

Along with color you can set: 

- `alpha=` the transparency (0 to 1)
- `size=` the size of point (0 and higher)
- `fill=` the color of the fill (doesn't apply to everything)

And more 


**Instructions**:

Start with your scatter plot with `cig_tax` on the y axis and `female_leg` on the x and then set the color to the `year` variable, but make all dots a little transparent by setting `alpha=0.5`. 


```{r aes, exercise=TRUE}
library(ggplot2)
ggplot(data=df) + geom_point(mapping=aes(x=female_leg, y=cig_tax))

```

```{r aes-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

ggplot(data=df) + geom_point(mapping=aes(x=female_leg, y=cig_tax, 
                                         color=year), alpha=0.5)  


```

```{r aes-check, eval=FALSE}
ex() %>% {
  check_function(., "ggplot") %>% check_arg("data") %>% check_equal()
  check_function(., "geom_point") %>% check_arg("alpha") %>% check_equal()
  check_function(., "aes") %>% {
    check_arg(., "x") %>% check_equal(eval = FALSE) 
    check_arg(., "y") %>% check_equal(eval = FALSE) 
    check_arg(., "color") %>% check_equal(eval = FALSE) 
  }
}
```
