---
title: "Hypothesis Testing and Residuals"
description: ""
output:
  learnr::tutorial:
    theme: sandstone
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(testwhat)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
tutorial_options(exercise.checker = testwhat::testwhat_learnr)
df <- read.csv("https://www.kevinreuning.com/assets/state_cig.csv")

```



## Confidence Intervals

R has a simple function called `confint()` that will calculate the confidence interval of all the variables in your model. It defaults to a 95% confidence interval but you can change that using `level = ` in the function call. 

**Instructions**:

We are going to recreate the model we've been using looking at the relationship between cigarette taxes and the percentage of liberals in a state. Remember, cigarette taxes is the DV and percentage liberals in the state is the IV. 

Once you have the model estimated you will then calculate the confidence interval. 

```{r linear-2, exercise=TRUE}
names(df)

```

```{r linear-2-solution, eval=FALSE}
df <- read.csv("https://www.kevinreuning.com/assets/state_cig.csv")

names(df)

mod <- lm(cig_tax~liberal, data=df)
confint(mod)

```

```{r linear-2-check, eval=FALSE}
ex() %>% {
  check_function(., "lm" ) %>% check_arg("formula") %>% check_equal()
  check_function(., "confint")
}

```




## Plotting Residuals

In the last assignment we saw how to plot residuals using `plot(mod$residuals)`. This plots them on the Y axis with the observation number on the X axis. We also are interested in plotting our residuals as a function of the the predicted value. This can be accomplished by using `mod$fitted.values`.  

**Instructions**:

Still using the same model, look at the residuals. We want to do two things:
1) Look at the residuals just by observation number.
2) Look at the residuals compared to the predicted value. 

```{r resids, exercise=TRUE}
names(df)
mod <- lm(cig_tax~liberal, data=df)

```

```{r resids-solution, eval=FALSE}
df <- read.csv("https://www.kevinreuning.com/assets/state_cig.csv")

names(df)

mod <- lm(cig_tax~liberal, data=df)
plot(mod$residuals)
plot(y=mod$residuals, x=mod$fitted.values)
```

```{r resids-check, eval=FALSE}
ex() %>% {
  check_function(., "plot", index=1 ) %>% check_arg("x") %>% check_equal
  check_function(., "plot", index=2 ) %>% {
    check_arg(., "x") %>% check_equal
    check_arg(., "y") %>% check_equal
  }

}

```

















