---
title: "Bivariate Regression in R"
description: "Bivariate regression and examining residuals in R"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(testwhat)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
tutorial_options(exercise.checker = testwhat_learnr)

df <- POL306::state_cig

```

## A Note on Data

We are still using the same data as we did in the plots lesson. Below is a reminder of what is in it. 

Throughout this tutorial we will use a dataset that looks at healthcare spending, cigarette taxes, and a variety of political variables. The unit of analysis for it is the state-year, and the data comes from 2001 to 2014. The variables are:

- `cig_tax` is the tax (in dollars) on a packet of cigarettes. 
- `female_leg` is the percentage of women in the state legislature.
- `liberal` is the percentage of voters that identify as liberal.
- `dem_gov` is an indicator that is 1 if the governor is a Democrat and 0 if they are not.
- `healthspendpc` is the amount spent perperson on healthcare. 
- `state` is the name of the state
- `year` is the year. 

## Bivariate Linear Regression

The most important function you will use for estimating linear regression models is `lm()`. This will expect a formula and will output something called a model object that you will want to save so you can then manipulate it. 

For example, if you have a dataframe called `data.frame` and three variables named: `var1` and `var2`. If you want to estimate a linear regression with `var1` as the dependent variable and `var2` as the independent variable you would write:

```
mod <- lm(var1 ~ var2, data=data.frame)
```

This won't output anything but you can look at the results using `summary()` like this:

`summary(mod)`

**Instructions**:

We will start by just estimating a linear regression of the cigarette tax (the DV) as a function of the percent of liberals in a state. The variables are all in the `df` dataset and are: `cig_tax`, and `liberal`. 

You should save the linear model as `mod` (like in the example above) and then output the results using the `summary()` function.



```{r linear, exercise=TRUE}
names(df) 
```

```{r linear-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

names(df)

mod <- lm(cig_tax~liberal, data=df)
summary(mod)
```

```{r linear-check, eval=FALSE}
ex() %>% {
  check_function(., "lm" ) %>%
    check_arg("formula") %>% check_equal()
  check_function(., "summary")
}
```



## Linear Regression Extended

The reason that we save the `lm()` object is that there are a lot of things we can do with it after we are done. Here are a few things:

- `coef(mod)` will create a vector of just the coefficients, which can be useful if we want to do something to them. 
- `mod$residuals` is a vector of all the residuals for the regression. We could easily plot them with `plot(mod$residuals)` (we'll see why this is important later)
- `nobs(mod)` calculates the number of observations in the final dataset.
- `mod$fitted.values` is a vector of the predictions from the regression.

**Instructions**:

Using the `mod` we made in the last activity (I included the code here), do the following:

1. Get the number of the observations included in the model. 
2. Plot the residuals.


```{r linear-2, exercise=TRUE}
names(df)

mod <- lm(cig_tax~liberal, data=df)
summary(mod)
```

```{r linear-2-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

names(df)

mod <- lm(cig_tax~liberal, data=df)
summary(mod)

nobs(mod)
plot(mod$residuals)
```

```{r linear-2-check, eval=FALSE}
ex() %>% {
  check_function(., "plot")
  check_function(., "nobs" )
}

```



## Plotting Residuals vs Predicted

We previously saw how to plot residuals using `plot(mod$residuals)`. This plots them on the y-axis with the observation number on the x-axis. We also are interested in plotting our residuals as a function of the the predicted value. This can be accomplished by using `mod$fitted.values` as well as `mod$residuals` and the `plot()` function. Remember you want the y-axis to have the residuals (you can specify which variable goes on which axis using `y=` and `x=`) and the x-axis should have the predicted values.


**Instructions**:

Still using the same model, now look at the residuals compared to the predicted value. You will have to include both variables in the call to `plot()`. 

```{r resids, exercise=TRUE}
names(df)
mod <- lm(cig_tax~liberal, data=df)

```

```{r resids-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

names(df)

mod <- lm(cig_tax~liberal, data=df)
plot(y=mod$residuals, x=mod$fitted.values)
```

```{r resids-check, eval=FALSE}
ex() %>% {
  check_function(., "plot") %>% {
    check_arg(., "x") %>% check_equal
    check_arg(., "y") %>% check_equal
  }

}

```


## Plotting a Regression Line

You can add the regression line to a scatter plot by using the `abline()` function after you make a scatter plot. You just have to put the model object (`mod`) into the `abline()` function. I've recreated the regression you looked at before with a scatter plot. Now you have to add a regression line. 


```{r linear-plot, exercise=TRUE}
names(df)

mod <- lm(cig_tax~liberal, data=df)
plot(y=df$cig_tax, x=df$liberal)
```

```{r linear-plot-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

names(df)

mod <- lm(cig_tax~liberal, data=df)
plot(y=df$cig_tax, x=df$liberal)
abline(mod)
```

```{r linear-plot-check, eval=FALSE}
ex() %>% 
  check_function("abline")

```


