---
title: "Hypothesis Testing and Confidence Intervals"
description: "How to do a hypothesis test and confidence intervals"
output:
  learnr::tutorial:
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(testwhat)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
tutorial_options(exercise.checker = testwhat_learnr)
df <- POL306::state_cig

```

## Hypothesis Testing by Hand 

Hypothesis testing can be done several different ways in R. Remember, at its basic though we are using the t-stat and comparing it to a critical value. The `summary()` of our model output will display the t-stat (though it calls it a t value). We can calculate the exact critical value by hand using the `qt()` function and running: 

- 1-sided greater than or less than: `-qt(alpha, degrees)`
- 2-sided: `qt(alpha/2, degrees)`

**Instructions**
We are going to recreate the model we've been using looking at the relationship between cigarette taxes and the percentage of liberals in a state. Remember, cigarette taxes is the DV and percentage liberals in the state is the IV. Save the t-stat as a new variable `t_stat` (remember: `variable <- 1.24`)

After you've done that, calculate what the critical value you should use to test a one-sided greater than hypothesis with 498 degrees of freedom and an alpha value of 0.05. Save this as `crit_val`

Finally, to test if your t-stat is greater than your critical value (what it needs to be to reject the null for a greater than hypothesis test) run `t_stat > crit_val`


```{r hyptest, exercise=TRUE}
names(df)

mod <- lm(cig_tax~liberal, data=df)

```

```{r hyptest-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

names(df)

mod <- lm(cig_tax~liberal, data=df)

t_stat <- 9.658
crit_val <- qt(0.05, 498)
t_stat > crit_val

```

```{r hyptest-check, eval=FALSE}
ex() %>% {
  check_function(., "lm" ) %>% check_arg("formula") %>% check_equal()
  check_object(., "crit_val") %>% check_equal()
  check_object(., "t_stat") %>% check_expr("round(t_stat, 1)") %>% check_result() %>% check_equal(incorrect_msg="t_stat needs to be the same as what is shown in the output of summary() with a rounding tolerance of 1 decimal point")
  check_code(., c("t_stat > crit_val", 
                  "crit_val < t_stat"), fixed=T, 
             missing_msg = "Be sure to check if the t_stat is greater than the crit_val")
}

```





## Confidence Intervals

R has a simple function called `confint()` that will calculate the confidence interval of all the variables in your model. It defaults to a 95% confidence interval but you can change that using `level = ` in the function call. 

**Instructions**:
Using the model you estimated before you will calculate the confidence interval. 

```{r linear-2, exercise=TRUE}
names(df)

mod <- lm(cig_tax~liberal, data=df)

```

```{r linear-2-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

names(df)

mod <- lm(cig_tax~liberal, data=df)
confint(mod)

```

```{r linear-2-check, eval=FALSE}
ex() %>% {
  check_function(., "lm" ) %>% check_arg("formula") %>% check_equal()
  check_function(., "confint")
}

```
















