---
title: "Hypothesis Testing and Confidence Intervals"
description: "How to do a hypothesis test and confidence intervals"
output:
  learnr::tutorial:
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(testwhat)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
tutorial_options(exercise.checker = testwhat_learnr)
df <- POL306::state_cig

```

## Hypothesis Testing by Hand 

Hypothesis testing can be done several different ways in R. Remember, at its basic though we are using the t-stat and comparing it to a critical value. The `summary()` of our model output will display the t-stat (though it calls it a t value). We can calculate the exact critical value by hand using the `qt()` function and running: 

- 1-sided greater than or less than: `-qt(alpha, degrees)`
- 2-sided: `qt(alpha/2, degrees)`

**Instructions**
We are going to recreate the model we've been using looking at the relationship between cigarette taxes and the percentage of liberals in a state. Remember, cigarette taxes is the DV and percentage liberals in the state is the IV. Save the t-stat as a new variable `t_stat` (remember: `variable <- 1.24`)

After you've done that, calculate what the critical value you should use to test a one-sided greater than hypothesis with 498 degrees of freedom and an alpha value of 0.05. Save this as `crit_val`

Finally, to test if your t-stat is greater than your critical value (what it needs to be to reject the null for a greater than hypothesis test) run `t_stat > crit_val`


```{r hyptest, exercise=TRUE}
names(df)

mod <- lm(cig_tax~liberal, data=df)

```

```{r hyptest-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

names(df)

mod <- lm(cig_tax~liberal, data=df)

t_stat <- 9.658
crit_val <- qt(0.05, 498)
t_stat > crit_val

```

```{r hyptest-check, eval=FALSE}
ex() %>% {
  check_function(., "lm" ) %>% check_arg("formula") %>% check_equal()
  check_object(., "crit_val") %>% check_equal()
  check_object(., "t_stat") %>% check_expr("round(t_stat, 1)") %>% check_result() %>% check_equal(incorrect_msg="t_stat needs to be the same as what is shown in the output of summary() with a rounding tolerance of 1 decimal point")
  check_code(., c("t_stat > crit_val", 
                  "crit_val < t_stat"), fixed=T, 
             missing_msg = "Be sure to check if the t_stat is greater than the crit_val")
}

```



## Confidence Intervals

R has a simple function called `confint()` that will calculate the confidence interval of all the variables in your model. It defaults to a 95% confidence interval but you can change that using `level = ` in the function call. 

**Instructions**:
Using the model you estimated before you will calculate the confidence interval. 

```{r linear-2, exercise=TRUE}
names(df)

mod <- lm(cig_tax~liberal, data=df)

```

```{r linear-2-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

names(df)

mod <- lm(cig_tax~liberal, data=df)
confint(mod)

```

```{r linear-2-check, eval=FALSE}
ex() %>% {
  check_function(., "lm" ) %>% check_arg("formula") %>% check_equal()
  check_function(., "confint")
}

```




## Looking at Substantive Effects
We can used CIs to examine the substantive relationship between an independent variable and its dependent variable. The way we discussed in class was by calculating how a standard deviation change in the independent variable would change the dependent variable. To do this we multiple the coefficient or a 95% CI by the standard deviation of the independent variable. We can calculate the latter using `sd()`. 

Before multiplying, we can make things bit simpler. `confint()` calculates the confidence interval for both the intercept and the coefficient of our IV and returns it as a 2 by 2 matrix. The CI for the IV is in the second row. If we run `confint(mod)[2,]` we'll just get the second row. 

**Instructions**
Estimate the effect of a 1 standard deviation change in `liberal` on `cig_tax` and save that as the variable `eff`. You can then compare that to the standard deviation of `cig_tax`. 


```{r substantive, exercise=TRUE}
names(df)

mod <- lm(cig_tax~liberal, data=df)

```

```{r substantive-solution, eval=FALSE}
df <- POL306::state_cig ### Ignore this

names(df)

mod <- lm(cig_tax~liberal, data=df)
confint(mod)[2,]
eff <- sd(df$liberal, na.rm=T) * confint(mod)[2,]
eff
sd(df$cig_tax, na.rm=T)

```

```{r substantive-check, eval=FALSE}
ex() %>% {
  check_function(., "lm" ) %>% check_arg("formula") %>% check_equal()
  check_function(., "confint")
  check_object(., "eff") %>% check_equal()
  check_function(., "sd")
}

```


